### 2014-08-22 ECGG_RTSProblems/ECGG_RTSProblems.tex
Added files to start working on the ECGG papers

\documentclass{llncs}

\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{color}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{mathabx}
%\usepackage{hyperref}
%\usepackage{subfigure}
%\usepackage[colorinlistoftodos, textwidth=3.2cm, shadow]{todonotes}

\usepackage{algorithm}
\usepackage{fixltx2e}
\usepackage{algpseudocode}

\usepackage{multirow}

% To use Call inside another Call (algorithms)
\MakeRobust{\Call}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Title
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{RTS AI Problems and Techniques
%\thanks{Supported by ...}
}

\author{Santiago~Onta\~{n}\'{o}n\inst{1} \and 
		Gabriel~Synnaeve\inst{2} \and
		Alberto~Uriarte\inst{1} \and
		Florian~Richoux\inst{3} \and
		David~Churchill\inst{4} \and
		Mike~Preuss\inst{5}
		}

\institute{
	Computer Science Department at Drexel University, Philadelphia, PA, USA. \\
	\email{\{santi,albertouri\}@cs.drexel.edu}
\and
	Cognitive Science and Psycholinguistics (LSCP) of ENS Ulm, Paris, France. \\
	\email{gabriel.synnaeve@gmail.com}
\and
	Nantes Atlantic Computer Science Laboratory (LINA) of the Universit{\'e} de Nantes, France. \\
	\email{florian.richoux@univ-nantes.fr}
\and
	Computing Science Department of the University of Alberta, Edmonton, Canada. \\
	\email{cdavid@cs.ualberta.ca}
\and
	Department of Computer Science of Technische Universit{\"a}t Dortmund, Germany.
	\email{mike.preuss@cs.tu-dortmund.de}
}

\begin{document}

\maketitle

\section{Definition}

{\color{blue} to-do}

\bibliographystyle{splncs}
\bibliography{references}

\end{document}